---
title: "Untitled"
author: "Zi Lynn Ow"
date: "10/20/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(RSQLite)
library(DBI)
library(tidyverse)
library(dbplyr)
library(fuzzyjoin)
```

```{r}
if (file.exists("airline2.db"))
  file.remove("airline2.db")
```

```{r}
conn <- dbConnect(RSQLite::SQLite(), "airline2.db")
```

```{r}
ontime <- list.files(path = "/Users/zilynnow/Desktop/ST2195_ProgrammingforDataScience/Assignment/Practice_Assignment_03/ontime", pattern = "*.csv", full.names = TRUE) %>% 
lapply(read_csv) %>% bind_rows                                                      
ontime                                                     
```

```{r}
airports <- read.csv("/Users/zilynnow/Desktop/ST2195_ProgrammingforDataScience/Assignment/Practice_Assignment_03/dataverse_files/airports.csv", header = TRUE)

carriers <- read.csv("/Users/zilynnow/Desktop/ST2195_ProgrammingforDataScience/Assignment/Practice_Assignment_03/dataverse_files/carriers.csv", header = TRUE)

planes <- read.csv("/Users/zilynnow/Desktop/ST2195_ProgrammingforDataScience/Assignment/Practice_Assignment_03/dataverse_files/plane-data.csv", header = TRUE)
```

```{r}
dbWriteTable(conn,"ontime", ontime)
```

```{r}
dbWriteTable(conn,"planes", planes)
dbWriteTable(conn,"airports", airports)
dbWriteTable(conn,"carriers", carriers)
```

```{r}
dbListFields(conn,"airports", airports)
```

```{r}
dbListFields(conn,"carriers", carriers)
```

```{r}
dbListFields(conn,"planes", planes)
```

```{r}
dbListFields(conn,"ontime", ontime)
```

```{r}
ontime_db <- tbl(conn, "ontime")
airports_db <- tbl(conn, "airports")
carriers_db <- tbl(conn, "carriers")
planes_db <- tbl(conn,"planes")
```

Q1: Which plane model has the lowest associated average departure delay (excluding cancelled and diverted flights)?

```{r}
q1 <- dbGetQuery(conn, 
"SELECT model AS model, AVG(ontime.DepDelay) AS avg_DepDelay
FROM ontime INNER JOIN planes USING (tailnum)
WHERE ontime.DepDelay>0 & ontime.Diverted == 0 & ontime.Cancelled == 0
GROUP BY model
ORDER BY avg_DepDelay")
print(paste(q1[1, "model"], "has the lowest associated average departure delay."))
```

```{r}
q1b <- ontime_db %>% rename_all(tolower) %>% inner_join(planes_db, by = "tailnum", suffix =c(".ontime",".planes")) %>% filter(DepDelay>0) %>% filter(Diverted == 0) %>% filter(Cancelled == 0 )%>% group_by(model) %>% summarize(avg_DepDelay = mean(DepDelay, na.rm = TRUE)) %>% arrange(avg_DepDelay) 
print(head(q1b, 1))
```

```{r}
dbWriteTable(conn,"q1", q1)
```

Q2: Which city has the highest number of inbound flights (excluding cancelled flights?)

```{r}
q2 <- dbGetQuery(conn, 
"SELECT airports.city AS airport, COUNT(*) as total
FROM airports INNER JOIN ontime
ON ontime.Dest = airports.iata
WHERE ontime.Cancelled = 0 
GROUP BY airports.city
ORDER BY total DESC")
print(paste(q2[1, "airport"], "as the highest number of inbound flights (excluding cancelled flights)."))
```

```{r}
q2b <- inner_join(ontime_db, airports_db, by = c("Dest"="iata")) %>% filter(Cancelled == 0) %>% group_by(city) %>% 
  count(city) %>% arrange(desc(n)) 
print(head(q2b, 1))
```

```{r}
dbWriteTable(conn,"q2", q2)
```

Q3: Which carrier has the highest number of cancelled flights?

```{r}
q3 <- dbGetQuery(conn, 
"SELECT carriers.Description AS carrier, COUNT(*) as total
FROM carriers INNER JOIN ontime 
ON ontime.UniqueCarrier = carriers.Code
WHERE ontime.Cancelled = 1
GROUP BY carriers.Description
ORDER BY total DESC")
print(paste(q3[1, "carrier"], "has the highest number of cancelled flights."))
```

```{r}
q3b <- inner_join(ontime_db, carriers_db,by=c("UniqueCarrier"="Code")) %>% filter(Cancelled ==1) %>%  
group_by(Description) %>% summarize(total= sum(Cancelled, na.rm = TRUE)) %>% arrange(desc(total))
print(head(q3b, 1))
```

```{r}
dbWriteTable(conn,"q3", q3)
```

Q4: Which carrier has the highest number of cancelled flights, relative to their number of total flights?

```{r}
UniqueCarrierCount <- dbGetQuery(conn, 
"SELECT carriers.Description AS carrier, COUNT(*) as denominator
FROM carriers INNER JOIN ontime 
ON ontime.UniqueCarrier = carriers.Code
GROUP BY carriers.Description
ORDER BY denominator DESC")
UniqueCarrierCount
```

```{r}
dbWriteTable(conn,"UniqueCarrierCount", UniqueCarrierCount)
```

```{r}
q4 <- dbGetQuery(conn, 
"SELECT *, total* 1./denominator as ratio
FROM q3 INNER JOIN UniqueCarrierCount USING(carrier)
ORDER BY ratio DESC")
print(paste(q4[1, "carrier"], "has the highest number of cancelled flights, relative to their number of total flights."))
```

```{r}
UniqueCarrierCount2 <- inner_join(ontime_db, carriers_db,by=c("UniqueCarrier"="Code")) %>% 
group_by(Description) %>% summarize(denominator= count(Description)) %>% arrange(desc(denominator)) 
UniqueCarrierCount2
```

```{r}
q3_db <- tbl(conn,"q3")
UniqueCarrierCount_db <- tbl(conn,"UniqueCarrierCount")
```

```{r}
q4b <- mutate(inner_join(q3_db, UniqueCarrierCount_db, by = "carrier"), ratio = total*1./denominator) %>% arrange(desc(ratio)) 
print(head(q4b, 1))
```

```{r}
dbWriteTable(conn,"q4", q4)
```

```{r}
#dbDisconnect(conn)
```
